\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{DATA420-A2}
\author{david.ewing.nz }
\date{September 2025}

\begin{document}

\maketitle

% Assignment Method Overview
% For Overleaf/LaTeX

\section*{Assignment Method Overview}

\begin{enumerate}
  \item \textbf{Data Exploration}
    \begin{enumerate}
      \item Overview of dataset structure, sizes, formats, data types, storage.\\
      Use HDFS or Spark commands to list files, check file sizes, and inspect file headers. Summarise formats (CSV, TSV, ARFF), data types (numeric, string), and storage locations.
      \item Loading different dataset types.\\
      Use Spark's read methods to load small samples, infer schema, and validate by displaying the first few rows.
      \item Counting rows and comparing to unique songs.\\
      Use Spark's count() and distinct().count() to compare row counts and unique song counts.
    \end{enumerate}
  \item \textbf{Data Preprocessing}
    \begin{enumerate}
      \item Loading audio feature attribute names and types.\\
      Read attribute files as text, parse to extract names/types, and use to define Spark schemas.
      \item Mapping attribute types to Spark types.\\
      Create a mapping (e.g., ``NUMERIC'' $\rightarrow$ FloatType), build StructType schema programmatically.
      \item Evaluating attribute names as column names.\\
      Review names for clarity, length, and uniqueness; discuss pros/cons for modelling and merging.
      \item Systematic renaming for merging.\\
      Propose a naming convention (e.g., prefix with dataset name), use Spark renaming functions.
    \end{enumerate}
  \item \textbf{Audio Similarity \\& Classification}
    \begin{enumerate}
      \item Merging audio feature datasets and generating statistics.\\
      Load datasets, join on track/song ID, compute descriptive statistics and correlations.
      \item Loading and analysing genre dataset.\\
      Load genre data, count frequencies, visualise, and discuss class imbalance.
      \item Merging genres and features.\\
      Join on track/song ID to create a labelled dataset for classification.
    \end{enumerate}
  \item \textbf{Binary Classification}
    \begin{enumerate}
      \item Algorithm selection and justification.\\
      Discuss why logistic regression, random forest, and GBT are suitable; propose preprocessing.
      \item Creating a binary label for ``Electronic''.\\
      Map genre column to binary, calculate and report class balance.
      \item Splitting data and handling class balance.\\
      Use stratified sampling for 80/20 split; justify resampling.
      \item Training classifiers.\\
      Fit each model using Spark ML, document training process.
      \item Evaluating performance.\\
      Use test set to compute accuracy, precision, recall; present results in a table.
      \item Comparing algorithms and metrics.\\
      Discuss which model performs best and why; explain metric choices.
    \end{enumerate}
  \item \textbf{Multiclass Classification}
    \begin{enumerate}
      \item Logistic regression for multiclass.\\
      Explain one-vs-rest or multinomial logistic regression, note extra configuration if needed.
      \item Encoding genres as integer labels.\\
      Use StringIndexer or similar, recalculate class balance.
      \item Training and evaluating multiclass model.\\
      Train on multiclass labels, compute multiclass metrics, discuss impact of class imbalance.
    \end{enumerate}
  \item \textbf{Hyperparameter Tuning}
    \begin{enumerate}
      \item Hyperparameters for each algorithm.\\
      List key hyperparameters, discuss expected effects and alternatives.
      \item Cross-validation explanation.\\
      Describe k-fold cross-validation and its role in tuning.
      \item Tuning process and expected impact.\\
      Outline grid/random search, estimate potential improvement in metrics.
    \end{enumerate}
  \item \textbf{Song Recommendation (Collaborative Filtering)}
    \begin{enumerate}
      \item Dataset partitioning and caching.\\
      Assess dataset size, recommend repartitioning/caching for efficiency.
      \item Loading and analysing user-song data.\\
      Count unique users and songs, find most active user, calculate percentages.
      \item Additional descriptive statistics.\\
      Propose further stats (e.g., average plays per user/song).
      \item Visualising distributions.\\
      Plot song popularity and user activity distributions, describe observed patterns.
    \end{enumerate}
\end{enumerate}


\end{document}
